An Analysis of the Use of Centipawn Loss to Detect Cheating in Chess

### 2022-11-14 Manuscript/manuscript.tex
edited intro, added more content to data

@@

% insert data from lichess

\section*{Methods}
\addcontentsline{toc}{section}{Methods}
\label{sec:meth}

\section*{Simulation}
\addcontentsline{toc}{section}{Simulation}
\label{sec:sim}

>>>>

Measures of average centipawn loss can be used to evaluate a player's strength and, theoretically, one can compare ACL to one's ELO rating, a measure of one's relative skill level in zero-sum games such as chess, to observe how well they played with respect to this ranking. The contribution of this paper is to provide an analysis of this method, as a means to detect cheating in chess, by observing suspicious historic activity (i.e., an individual playing well above their rating in terms of ACL) with the goal of scrutinizing whether it should be treated as a "smoking gun" in terms of evidence of cheating; as was the case with Hans Niemann.

The rest of the paper is organized as follows. Section 2 provides specifics about the data used for the ACL versus ELO rating analysis, such as how it was collected and what it does to help answer my research question. Section 3 will provide insight into my methods and computer programs used to perform the analysis. Section 4 concludes with a discussion.

@@

The data used was provided by Lichess.com, a free and open-source internet chess server that holds the records of many professional games. It is a comprehensive list of every professional game Hans Niemann has played, both virtually and OTB, from March of 2018 to September of 2022 and is provided in both .csv and .pkl format. Some of the relevant information it contains are what color Niemann and his opponent played and what their respective ELO scores and calculated ACL scores were along with some other unimportant details for this analysis (e.g., ACL score of Niemann's opponent, tournament name and event round, etc.). The main information of concern with regard to the entirety of the data is Niemann's ELO score and calculated ACL score of each game of each particular point in time, as this is what will be used to plot. All other games that included information where Niemann was not playing were dropped from the original dataset provided by Lichess.

\section*{Simulation/Methods}
\addcontentsline{toc}{section}{Methods}
\label{sec:sim}
